The Werewolf Among Us: Humans vs LLMs in Multi-Agent Games

### 2025-05-06 _tex/paper.tex
Built site for gh-pages

\begin{itemize}
\item
  Social deduction games like \emph{Werewolf} offer a clear way to
  evaluate how agents deceive, persuade, and reason in group
  settings(\citeproc{ref-wikiwerewolf}{Wikipedia contributors 2024}). In
  these games, players have limited information, hidden identities, and
  must convince others while trying to figure out who is lying. These
  challenges closely match real life situations involving trust,
  negotiation, and manipulation.
\item
  We wanted to compare how humans and large language models (LLMs)
  handle these situations. To do this, we used two recent datasets:

@@

\begin{itemize}
\tightlist
\item
  Human dataset description
  (\citeproc{ref-laiWerewolfUsMultimodal2022}{Lai et al. 2022})
\item
  Is specifically for a form of one-night werewolf

  \begin{itemize}
  \tightlist
  \item
    Describe key differences
  \end{itemize}
\item
  Used specifically for the text available

  \begin{itemize}
  \tightlist
  \item
    and annotations of persuasion strategy on the text
  \end{itemize}
\end{itemize}

\subsection{Werewolf Arena}\label{werewolf-arena}

\begin{itemize}
\tightlist
\item
  (\citeproc{ref-bailisWerewolfArenaCase2024}{Bailis, Friedhoff, and
  Chen 2024})
\item
  Discuss the framework, how it works, prompts, etc
\item
  Discuss what types of runs we did
\item
  Discuss the data included in output
\item
  Talk about how we had to annotate the LLM speech with persuasion
  strategies ourselves
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Formatted data to match, performed various comparisons
\end{itemize}

>>>>

\begin{itemize}
\item
  Social deduction games like \emph{Werewolf} offer a clear way to
  evaluate how agents deceive, persuade, and reason in group
  settings(\citeproc{ref-wikiwerewolf}{Wikipedia contributors 2024}). In
  these games, players have limited information, hidden identities, and
  must convince others while trying to figure out who is lying. These
  challenges closely match real life situations involving trust,
  negotiation, and manipulation.
\item
  We wanted to compare how humans and large language models (LLMs)
  handle these situations. To do this, we used a recent human dataset
  and constructed one from an LLM-based simulator:

@@

We used the \emph{Werewolf Among Us} dataset (Lai et al., 2022), which
contains annotated dialogues from over 150 real games of One Night
Werewolf and Avalon. These games differ from classic Werewolf because
they have only one round of discussion and voting, do not eliminate
players during the game, and include many specialized roles beyond
Villager and Werewolf. The dataset provides detailed annotations of
persuasion strategies for each utterance, such as accusations, defenses,
and identity claims. We specifically used the textual transcriptions and
their strategy annotations for our comparisons.

\subsection{Werewolf Arena (LLM
Dataset)}\label{werewolf-arena-llm-dataset}

The \emph{Werewolf Arena} dataset (Bailis et al., 2024) features
simulated classic Werewolf games played by LLM agents without human
intervention. Unlike the one-round human games, these simulations
involve multiple rounds alternating between night (secret actions) and
day (open discussion). Each LLM agent is assigned a role (Villager,
Werewolf, Seer, Doctor) and receives context-specific prompts that guide
their strategic interactions.

We ran simulations using five LLM models: GPT-4o, GPT-4.1, GPT-4o-mini,
DeepSeek-Chat, and DeepSeek-Reasoner. We experimented with two
configurations:

\begin{itemize}
\item
  8 players with 8 rounds of discussion
\item
  10 players with 6 rounds of discussion
\end{itemize}

We chose these configurations to give villagers more time to coordinate,
extending the dialogue to better observe persuasive behaviors. After
running the simulations, we manually annotated the LLM-generated
dialogue using the same persuasion categories as in the human dataset.

@@

We standardized annotations across both datasets so they could be
directly compared. Our analyses focused on frequency distributions of
persuasion strategies, comparisons by role (villager versus werewolf),
and differences in how humans and LLMs applied these strategies.
